On the relation between metacognitive monitoring and domain knowledge

### 2019-12-12 ms.Rmd
Submitted to BRM

abstract: |
  Metacognitive resolution refers to people's ability to tell their correct answers apart from incorrect answers, and is theorized an important factor in learning. Resolution---the trial-by-trial association between response accuracy and metacognitive endorsement of that response's accuracy---is a distinct ability from knowledge in the domain of testing, and instead indicates a higher-order self-evaluation of one's cognitive states. It is therefore important that measures of resolution are independent of domain knowledge. We conducted six experiments that revealed a robust positive correlation between metacognitive resolution and performance in multiple-choice mathematics testing. A simulation study suggested that resolution metrics are increasingly negatively biased with decreasing performance, because multiple-choice tasks permit correct guessing. Finally, an experiment supported the guessing-related confound hypothesis: Resolution and performance were positively correlated in multiple-choice testing, but not in free-response testing. This study brings to light a previously underappreciated limitation in assessing metacognitive resolution and its relation to task performance when the criterion task involves guessing.

@@

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
source("functions.R")
library(papaja)
library(knitr)
library(Hmisc)
library(afex)
library(ggbeeswarm)
library(flextable)
library(patchwork)
library(broom)
library(scales)
library(tidyverse)
library(glue)
library(ordinal)
library(beepr)
library(parallel)
opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  cache = TRUE,
  dev = c("png", "pdf"),
  fig.ext = c("png", "pdf"),
  fig.retina = 2
)
options(mc.cores = parallel::detectCores())
theme_set(
  theme_linedraw(base_size = 8) + 
    theme(panel.grid = element_blank(), legend.position = "right")
)
# Function to format p values
pvalue <- function(x) {
  out <- scales::pvalue(x, add_p = T)
  out <- str_replace(out, "<", " < ")
  out <- str_replace(out, "=", " = ")
  out
}
```

@@

Metacognitive resolution is typically measured with metrics that are asserted to be independent of the participants' performance, skill, or knowledge in the primary task. For example, in a memory task, participants' memory abilities should not impact their metacognitive abilities, and vice versa. In an influential article @Nelsoncomparisoncurrentmeasures1984 recommended the Goodman-Kruskal gamma correlation [@GoodmanMeasuresAssociationCross1954] as a measure of resolution, partly because it is theoretically independent of task performance [see also @GonzalezMeasuringordinalassociation1996]. Other measures of resolution include Pearson's *r* [@NelsonNorms300generalinformation1980], the phi correlation [@KornellTransferMetacognitiveSkills2007], and signal detection theoretic measures [@benjamin2008measurement; @HighamNewimprovedgamma2018]. Although computationally different, these metrics measure the association between response accuracy and metacognitive rating in a way that is usually assumed to quantify metacognitive ability independent of task performance. However, it remains a possibility that metacognitive resolution is empirically related to task performance, with important implications to self-directed study.

@@

In what we call Experiment 1, we collated data from six experiments investigating mathematics learning. There were differences between experiments, such as testing medium (computer or paper) and age (middle school or undergraduate), but our focus here was not to investigate these factors' impact on mathematics performance or metacognition. Rather, we used this large dataset to examine the possible relationship between metacognitive resolution and (mathematics) performance, in a multiple-choice testing context. The common procedure was that participants completed mathematics tests consisting of 4-alternative multiple-choice problems, and rated their confidence in the accuracy of each of their answers.

With respect to resolution in mathematics study, two previous studies which contained a resolution measure have reported conflicting results. @SieglerDevelopmentalindividualdifferences2013, in a study of middle school students' fraction arithmetic skills, reported that students' confidence ratings were on average 0.7 points greater for correct than incorrect answers on a 1-5 scale, suggesting that young learners of mathematics have reasonably accurate metacognitions about their answers' accuracy (though the difference in confidence is small). @SchrawEffectGeneralizedMetacognitive1996 tested university undergraduates on probability problems and had them rate their confidence in their answers. The correlation between confidence and accuracy in that study was not statistically significant (Pearson's *r* = -.03)[^1], indicating that university students' metacognitive resolution (within probability calculations) is not accurate. However, neither study addressed the potential correlation between resolution and mathematics performance, which we investigate here.

[^1]: The description of how this correlation was calculated is not clear to us, but we take it that the coefficient reflects a cross-participant mean of within-participant Pearson correlations. @SchrawEffectGeneralizedMetacognitive1996 also reported a correlation in which the data for all participants was pooled, and this correlation was positive (*r* = .16) but not statistically significant.

Because of the similarity of these experiments (1A-F), we describe all six experiments' methods first, and then present the results together.

@@

```{r math-example, out.width = "4cm", fig.cap = "Example mathematics problem from Experiment 1."}
include_graphics("example-problem-14.8.13.png")
```

@@

Experiment 1C was a close but not exact replication of Experiments 1A and B. Ninety-one 8^th^ grade students participated from the same New York City public school as in Experiments 1A and B. We obtained voluntary consent from the students and their parents. The experiment was approved by the Columbia University Internal Review Board and by the New York City Board of Education.

The students completed three pen-and-paper tests, separated by eight days of in-class instruction, as part of an after-school mathematics tutorial. Each test consisted of 20 four-alternative multiple-choice problems from the same pool of problems as in experiments 1A and B. Multiple problems were printed on each page of the test, but otherwise the test procedure was the same as in experiments 1A and B: Participants wrote their answers and confidence ratings---on the same scale as in the previous experiments---on a grid on the first page of the test. Participants completed the tests in groups of up to 25 students in a classroom supervised by a teacher, and had 45 minutes for each test.

@@

Experiment 1D was similar to previous experiments, but the tests were completed on a computer. Eighty-nine students participated. The participants were the same students as those in Experiment 1C (two students who completed Experiment 1C did not take part in Experiment 1D).

In Experiment 1C, the participants had completed three pen-and-paper tests, separated by eight weeks of instruction. During those eight weeks, the participants engaged in eight computerized mathematics tutorials, once per week. Although the same students participated, the procedure was different to Experiment 1C and we therefore described it as Experiment 1D.

@@

Eighty-six Columbia University undergraduate students participated for course credit in an introductory Psychology course (36 females, 37 males, the rest did not report gender; age: 18-36 years). The experiment was approved by the Columbia University Internal Review Board.

The participants completed two tests, each consisting of 20 math problems. Each problem, and its response options, was printed on a separate page. Participants were asked to solve the problems and write their preferred response alternative and confidence rating on a separate answer sheet. Participants provided confidence ratings by writing a number from 0 ("low") to 10 ("high") next to their chosen response on the answer sheet. The two tests were separated by a computerized mathematics tutorial, which included videos about solving problems related to the ones on the tests, which we did not consider here. The problems were selected from the same pool of materials as in the previous experiments. Up to four individuals participated in the same room, under experimenter supervision. Participation took up to 90 minutes, and participants were allowed scrap paper and calculators.

@@

Experiment 1F was similar to Experiment 1E, but the mathematics tests were computerized. Ninety-nine undergraduates participated for course credit. The experiment was approved by the Columbia University Internal Review Board.

In this experiment, participants completed four mathematics tests as part of an online mathematics tutorial. Each test consisted of ten problems, presented alone on the top of the screen, with the answer alternatives below. Participants chose their preferred response by clicking a box next to it. They then clicked "Submit", and proceeded to the next screen, where they rated their confidence in the accuracy of their response by clicking on a visual analog scale. After indicating their degree of confidence, they clicked "Submit" to proceed to the next screen. The following screen allowed the participant to view different types of instructional videos related to the mathematics problems, but these are not analyzed in the current study. The tests were completed in one 90-minute session, on individual computers, in groups of up to four supervised by an experimenter.

@@

The mean gamma correlations and *d*~a~s for Experiments 1A-F are summarized in Table \@ref(tab:math-dtab). As can be seen mathematics performance was close to 70% correct across experiments. More importantly, metacognitive resolution was good and significantly better than chance in all six experiments.

```{r math-dtab}
new_names <- tibble(
  key = c("zroc_auc_gamma", "gamma_old", "zroc_da", "phi", "zroc_auc", "trap_auc", "proportion_correct", "mean_rating"),
  Metric = c("Gamma (ROC)", "Gamma (pairs)", "da", "Phi", "AUC (zROC)", "AUC (trap)", "Proportion Correct", "Mean confidence")
)
tab <- math_metrics %>%
  gather(key, value, proportion_correct, zroc_auc_gamma, zroc_da, gamma_old) %>%
  group_by(key, Experiment) %>%
  group_modify(~ data.frame(t(smean.cl.boot(.x$value, B = 1000)))) %>%
  ungroup() %>%
  left_join(new_names) %>%
  right_join(math_subjects) %>%
  mutate_at(vars(Mean, Lower, Upper), ~ format(round(., 2), nsmall = 2)) %>%
  mutate(result = glue("{Mean} [{Lower}, {Upper}]")) %>%
  select(Experiment, N, Metric, result) %>%
  spread(Metric, result) %>%
  select(Experiment, N, `Proportion Correct`, everything())
kable(tab, caption = "Measures of mathematics performance and resolution in Experiment 1. Numbers indicate means with 95\\% CIs in square brackets.")
```

@@

```{r edge-gammas, eval = FALSE}
math_metrics %>% 
  filter(gamma_old == 1) %>% 
  left_join(math) %>% 
  ggplot(aes(con, fill = factor(accuracy))) +
  geom_histogram(position = "dodge") +
  facet_wrap("subject", scales = "free")
```

@@

```{r math-rtab}
tab <- math_metrics %>%
  gather(key, value, zroc_da, zroc_auc_gamma, gamma_old) %>%
  left_join(new_names) %>%
  group_by(Experiment, Metric) %>% 
  group_modify(
    ~tidy(cor.test(.x$value, .x$proportion_correct, method = "kendall"))
  ) %>%
  mutate(p.value = pvalue(p.value)) %>%
  mutate(result = format(round(estimate, 2), nsmall = 2, trim = TRUE)) %>%
  mutate(result = glue("{result} ({p.value})")) %>%
  select(Experiment, Metric, result) %>%
  ungroup() %>%
  pivot_wider(names_from = c(Metric), values_from = result)
kable(
  tab, 
  caption = "Correlation between mathematics performance and three metrics of metacognitive resolution in Experiment 1."
)
```

@@

```{r sim-math-perfect}
tmp1 <- math_metrics %>%
  select(Experiment, subject, proportion_correct, zroc_da_data = zroc_da) %>% 
  mutate(proportion_correct = plyr::round_any(proportion_correct, .1, floor))
tmp2 <- sims_mean %>% 
  filter(d_gen == max(delta_bins), p_guess == "0.25") %>% 
  select(proportion_correct, zroc_da_sim = zroc_da) %>% 
  mutate(proportion_correct = plyr::round_any(proportion_correct, .1, floor)) %>% 
  distinct(proportion_correct, .keep_all = T)
tmp <- left_join(tmp1, tmp2) %>% 
  mutate(diff = zroc_da_data - zroc_da_sim)
# Are subjects gamma/da significantly lower than the gamma/da from simulation with perfect metacognition? That is, we use the difference between each subject's gamma and the simulated perfect gamma at that subjects point of performance
res <- papaja::apa_print.htest(t.test(tmp$diff))$statistic
```

Finally, we asked whether the participants' resolution was different from perfect if guessing was accounted for. We tested participants' *d*~a~s against the estimates from the simulation with Guessing = 0.25 and Resolution as the greatest value used in our simulations (2.5), at the level of the participants' performance, with a paired *t*-test. The simulated mean estimates of *d*~a~ from near-perfect resolution and Guessing = 0.25 were greater than the participants' observed *d*~a~s (`r res`): Participants did not have perfect metacognitive resolution. Further, better performing participants were further below near-perfect resolution than were poorer performing participants. Using gamma as the resolution metric yielded identical results.

@@

In summary, our simulation study suggested a general and important source of variation in resolution metrics in studies of metacognition using multiple-choice testing: The more opportunities there are for correct guesses---either due to number of response alternatives or to knowledge on the primary task---the more will measures of resolution be biased toward being too small. However, in order to examine this bias, we assumed a generative model of confidence ratings that relied on random guessing in the primary task. One concern with this model was that such random guessing, while intuitively appealing, may not be empirically plausible [e.g. @HighamNoSpecialsignal2007].

Therefore, we put the idea that resolution metrics are related to domain knowledge in multiple-choice tasks, possibly due to guessing, to an experimental test. In Experiment 2, we directly compared indices of metacognitive resolution, and their relation to domain skill, between free-response trials and multiple-choice trials.

@@

```{r trivia-dtab}
tmp1 <- trivia_metrics %>% 
  select(
    session, condition, proportion_correct, 
    mean_rating, gamma_old, zroc_da, zroc_auc_gamma
  ) %>% 
  gather(key, value, -session, -condition) %>% 
  spread(condition, value) %>% 
  mutate(difference = `2AFC`-Free) %>% 
  group_by(key) %>% 
  group_modify(~tidy(t.test(.$difference))) %>% 
  # mutate(p.value = pvalue(p.value), parameter = round(parameter, 0)) %>% 
  mutate_if(is.numeric, ~format(round(., 2), nsmall = 2)) %>%
  ungroup() %>% 
  transmute(
    key = key,
    result = glue("{estimate} [{conf.low}, {conf.high}]")
  ) %>% 
  left_join(new_names) %>% 
  select(-key) %>% 
  spread(Metric, result) %>% 
  add_column(condition = "2AFC - Free", .before = 1)
tmp2 <- trivia_metrics %>%
  # left_join(tmp_mc) %>% 
  gather(
    key, 
    value, 
    proportion_correct, mean_rating, gamma_old, zroc_auc_gamma, zroc_da,
  ) %>% 
  group_by(key, condition) %>%
  group_modify(~ data.frame(t(smean.cl.boot(.x$value, B = 1000)))) %>%
  ungroup() %>%
  left_join(new_names) %>%
  # mutate(Experiment = factor(exp)) %>%
  mutate_at(vars(Mean, Lower, Upper), ~ format(round(., 2), nsmall = 2)) %>%
  mutate(result = glue("{Mean} [{Lower}, {Upper}]")) %>%
  select(condition, Metric, result) %>%
  spread(Metric, result) %>%
  select(condition, `Proportion Correct`, `Mean confidence`, everything())
# ft <- flextable(tab)
# ft <- set_caption(tab, "Table 1. Descriptives of Study 1.")
# ft
bind_rows(tmp2, tmp1) %>% 
  kable(caption = "Measures of general information test performance and metacognition in Experiment 2. Numbers indicate means with 95\\% CIs in square brackets.")
```

@@

```{r trivia-ctab}
tmp <- trivia_metrics %>%
  gather(key, value, gamma_old, zroc_da, zroc_auc_gamma) %>%
  left_join(new_names) %>%
  group_by(condition, Metric)
tab <- group_modify(
  tmp,
  ~tidy(cor.test(.x$value, .x$proportion_correct, method = "kendall"))
) %>%
  mutate(p.value = pvalue(p.value)) %>%
  mutate(result = format(round(estimate, 2), nsmall = 2, trim = TRUE)) %>%
  mutate(result = glue("{result} ({p.value})")) %>%
  select(condition, Metric, result) %>%
  ungroup() %>%
  pivot_wider(names_from = c(Metric), values_from = result)
kable(
  tab, 
  caption = "Correlation between general information test performance and three metrics of metacognitive resolution in Experiment 2."
  )
```

```{r trivia-rfig, fig.cap = "Scatterplot of participants' resolution metrics and proportions correct across the two conditions in Experiment 2. Lines of best fit and 95\\% CIs are from an ordinary linear regression model.", fig.height = 3}
trivia_metrics %>%
  gather(key, value, zroc_auc_gamma, zroc_da, gamma_old) %>%
  left_join(new_names) %>% 
  ggplot(aes(proportion_correct, value)) +
  geom_hline(yintercept = 0, lty = 2, size = .25) +
  scale_y_continuous(
    "Resolution",
    breaks = pretty_breaks()
  ) +
  scale_x_continuous(
    "Proportion correct",
    breaks = pretty_breaks()
  ) +
  geom_point(shape = 21, alpha = .5, size = .75) +
  geom_smooth(method = "lm", col = "black", size = .4) +
  facet_grid(Metric ~ condition, scales = "free_y") +
  theme(aspect.ratio = 1)
```

>>>>

abstract: |
  Metacognitive resolution refers to people's ability to tell their correct answers apart from incorrect answers, and is theorized an important factor in learning. Resolution---the trial-by-trial association between response accuracy and metacognitive endorsement of that response's accuracy---is a distinct ability from knowledge in the domain of testing, and instead indicates a higher-order self-evaluation of one's cognitive states. It is therefore important that measures of resolution are independent of domain knowledge. We conducted six experiments that revealed a robust positive correlation between metacognitive resolution and performance in multiple-choice mathematics testing. A simulation study suggested that resolution metrics are increasingly negatively biased with decreasing performance, because multiple-choice tasks permit correct guessing. Therefore, the observed positive correlation was likely an artifact rather than a true correlation between psychological abilities. Finally, an experiment supported the guessing-related confound hypothesis: Resolution and performance were positively correlated in multiple-choice testing, but not in free-response testing. This study brings to light a previously underappreciated limitation in assessing metacognitive resolution and its relation to task performance when the criterion task involves guessing.

@@

documentclass     : "apa6"
classoption       : "man"
output: 
  papaja::apa6_pdf:
    keep_tex: false
---

```{r setup, include = FALSE}
source("functions.R")
library(papaja)
library(knitr)
library(Hmisc)
library(afex)
library(ggbeeswarm)
library(flextable)
library(patchwork)
library(broom)
library(scales)
library(tidyverse)
library(glue)
library(ordinal)
library(beepr)
library(parallel)
opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  cache = TRUE,
  dev = c("png", "pdf"),
  fig.ext = c("png", "pdf"), 
  fig.align = "center",
  fig.retina = 2
)
options(mc.cores = parallel::detectCores())
theme_set(
  theme_linedraw(base_size = 8) + 
    theme(panel.grid = element_blank(), legend.position = "right")
)
# Function to format p values
pvalue <- function(x) {
  out <- scales::pvalue(x, add_p = T)
  out <- str_replace(out, "<", " < ")
  out <- str_replace(out, "=", " = ")
  out
}
```

@@

Metacognitive resolution is typically measured with metrics that are asserted to be independent of the participants' performance, skill, or knowledge in the primary task. For example, in a memory task, participants' memory abilities should not impact their metacognitive abilities, and vice versa. In an influential article @Nelsoncomparisoncurrentmeasures1984 recommended the Goodman-Kruskal gamma correlation [@GoodmanMeasuresAssociationCross1954] as a measure of resolution, partly because it is theoretically independent of task performance [see also @GonzalezMeasuringordinalassociation1996]. Other measures of resolution include Pearson's *r* [@NelsonNorms300generalinformation1980], the phi correlation [@KornellTransferMetacognitiveSkills2007], and signal detection theoretic measures [@benjamin2008measurement; @HighamNewimprovedgamma2018]. Although computationally different, these metrics measure the association between response accuracy and metacognitive rating in a way that is assumed to quantify metacognitive ability independent of task performance. However, it remains a possibility that metacognitive resolution is empirically related to task performance, with important implications to self-directed study.

@@

In what we call Experiment 1, we collated data from six experiments investigating mathematics learning. There were differences between experiments, such as testing medium (computer or paper) and age (middle school or undergraduate), but our focus here was not to investigate these factors' impact on mathematics performance or metacognition. Rather, we used this large dataset to examine the possible relationship between metacognitive resolution and (mathematics) performance, in a multiple-choice testing context. The common procedure was that participants completed mathematics tests consisting of 4-alternative multiple-choice problems, and rated their confidence in the accuracy of each of their answers. Because these experiments (1A-F) were so similar, we describe all six experiments' methods first, and then present the results together.

@@

```{r math-example, out.width = "12cm", fig.cap = "Example mathematics problem from Experiment 1."}
include_graphics("example-problem-14.8.13.png")
```

@@

Experiment 1C was a close but not exact replication of Experiments 1A and B. Ninety-one 8^th^ grade students participated from the same New York City public school as in Experiments 1A and B. We obtained voluntary consent from the students and their parents. The experiment was approved by the Columbia University Internal Review Board and by the New York City Board of Education. The students completed three pen-and-paper tests, separated by eight days of in-class instruction, as part of an after-school mathematics tutorial. Each test consisted of 20 four-alternative multiple-choice problems from the same pool of problems as in experiments 1A and B. Multiple problems were printed on each page of the test, but otherwise the test procedure was the same as in experiments 1A and B: Participants wrote their answers and confidence ratings---on the same scale as in the previous experiments---on a grid on the first page of the test. Participants completed the tests in groups of up to 25 students in a classroom supervised by a teacher, and had 45 minutes for each test.

@@

Experiment 1D was similar to previous experiments, but the tests were completed on a computer. Eighty-nine students participated. The participants were the same students as those in Experiment 1C (two students who completed Experiment 1C did not take part in Experiment 1D). In Experiment 1C, the participants had completed three pen-and-paper tests, separated by eight weeks of instruction. During those eight weeks, the participants engaged in eight computerized mathematics tutorials, once per week. Although the same students participated, the procedure was different to Experiment 1C and we therefore described it as Experiment 1D.

@@

Eighty-six Columbia University undergraduate students participated for course credit in an introductory Psychology course (36 females, 37 males, the rest did not report gender; age: 18-36 years). The experiment was approved by the Columbia University Internal Review Board. The participants completed two tests, each consisting of 20 math problems. Each problem, and its response options, was printed on a separate page. Participants were asked to solve the problems and write their preferred response alternative and confidence rating on a separate answer sheet. Participants provided confidence ratings by writing a number from 0 ("low") to 10 ("high") next to their chosen response on the answer sheet. The two tests were separated by a computerized mathematics tutorial, which included videos about solving problems related to the ones on the tests, which we did not consider here. The problems were selected from the same pool of materials as in the previous experiments. Up to four individuals participated in the same room, under experimenter supervision. Participation took up to 90 minutes, and participants were allowed scrap paper and calculators.

@@

Experiment 1F was similar to Experiment 1E, but the mathematics tests were computerized. Ninety-nine undergraduates participated for course credit. The experiment was approved by the Columbia University Internal Review Board. Participants completed four mathematics tests as part of an online mathematics tutorial. Each test consisted of ten problems, presented alone on the top of the screen, with the answer alternatives below. Participants chose their preferred response by clicking a box next to it. They then clicked "Submit", and proceeded to the next screen, where they rated their confidence in the accuracy of their response by clicking on a visual analog scale. After indicating their degree of confidence, they clicked "Submit" to proceed to the next screen. The following screen allowed the participant to view different types of instructional videos related to the mathematics problems, but these are not analyzed in the current study. The tests were completed in one 90-minute session, on individual computers, in groups of up to four supervised by an experimenter.

@@

The mean gamma correlations and *d*~a~s for Experiments 1A-F are summarized in Table \@ref(tab:math-dtab). Mathematics performance was close to 70% correct across experiments. More importantly, metacognitive resolution was good and significantly better than chance in all six experiments.

```{r math-dtab, results = 'asis'}
new_names <- tibble(
  key = c("zroc_auc_gamma", "gamma_old", "zroc_da", "phi", "zroc_auc", "trap_auc", "proportion_correct", "mean_rating"),
  Metric = c("Gamma (ROC)", "Gamma (pairs)", "da", "Phi", "AUC (zROC)", "AUC (trap)", "Proportion Correct", "Mean confidence")
)
tab <- math_metrics %>%
  gather(key, value, proportion_correct, zroc_auc_gamma, zroc_da, gamma_old) %>%
  group_by(key, Experiment) %>%
  group_modify(~ data.frame(t(smean.cl.boot(.x$value, B = 1000)))) %>%
  ungroup() %>%
  left_join(new_names) %>%
  right_join(math_subjects) %>%
  mutate_at(vars(Mean, Lower, Upper), ~ format(round(., 2), nsmall = 2)) %>%
  mutate(result = glue("{Mean} [{Lower}, {Upper}]")) %>%
  select(Experiment, N, Metric, result) %>%
  spread(Metric, result) %>%
  select(Experiment, N, `Proportion Correct`, everything())
kable(
  tab,
  booktabs = TRUE,
  escape = FALSE,
  caption = "Measures of mathematics performance and resolution in Experiment 1. Numbers indicate means with 95\\% CIs in square brackets."
)
```

@@

@@

```{r math-rtab, results = 'asis'}
tab <- math_metrics %>%
  gather(key, value, zroc_da, zroc_auc_gamma, gamma_old) %>%
  left_join(new_names) %>%
  group_by(Experiment, Metric) %>% 
  group_modify(
    ~tidy(cor.test(.x$value, .x$proportion_correct, method = "kendall"))
  ) %>%
  mutate(p.value = pvalue(p.value)) %>%
  mutate(result = format(round(estimate, 2), nsmall = 2, trim = TRUE)) %>%
  mutate(result = glue("{result} ({p.value})")) %>%
  select(Experiment, Metric, result) %>%
  ungroup() %>%
  pivot_wider(names_from = c(Metric), values_from = result)
papaja::apa_table(
  tab, 
  caption = "Correlation between mathematics performance and three metrics of metacognitive resolution in Experiment 1."
)
```

@@

```{r sim-math-perfect, eval = FALSE}
tmp1 <- math_metrics %>%
  select(Experiment, subject, proportion_correct, zroc_da_data = zroc_da) %>% 
  mutate(proportion_correct = plyr::round_any(proportion_correct, .1, floor))
tmp2 <- sims_mean %>% 
  filter(d_gen == max(delta_bins), p_guess == "0.25") %>% 
  select(proportion_correct, zroc_da_sim = zroc_da) %>% 
  mutate(proportion_correct = plyr::round_any(proportion_correct, .1, floor)) %>% 
  distinct(proportion_correct, .keep_all = T)
tmp <- left_join(tmp1, tmp2) %>% 
  mutate(diff = zroc_da_data - zroc_da_sim)
# Are subjects gamma/da significantly lower than the gamma/da from simulation with perfect metacognition? That is, we use the difference between each subject's gamma and the simulated perfect gamma at that subjects point of performance
res <- papaja::apa_print.htest(t.test(tmp$diff))$statistic
```

@@

In summary, our simulation study suggested that in studies of metacognition using multiple-choice testing, metrics of resolution will necessarily correlate with primary task performance. However, we assumed a generative model of confidence that relied on random guessing in the primary task. One concern with this model is that such random guessing, while intuitively appealing, may not be empirically plausible [e.g. @HighamNoSpecialsignal2007]. In fact, one study using SAT (five alternative multiple-choice) materials found that a signal detection theoretic (SDT) model described performance better than did a high-threshold model (i.e. model with random guessing; @HighamNoSpecialsignal2007). Specifically, high-threshold theory (as implied in our simulations) predicts that accuracy at the lowest level of confidence should not drop below 1/K (K = number of response alternatives). In contrast, as found in that study and predicted by SDT, accuracy was below 20% for low confidence responses, thus discrediting the random guessing assumption [@HighamNoSpecialsignal2007].

```{r accuracy-at-confidence, eval = TRUE}
tmp <- bind_rows(
  math %>% 
    filter(Experiment %in% c("1A", "1B", "1C")) %>% 
    group_by(Experiment, subject) %>% 
    mutate(con = bin_ratings(con, 1, 6, 6)) %>% 
    group_by(Experiment, subject, con) %>% 
    summarise(p = mean(accuracy)),
  math %>% 
    filter(Experiment %in% c("1D", "1F")) %>% 
    group_by(Experiment, subject) %>% 
    mutate(con = bin_ratings(con, 0, 1, 6)) %>% 
    group_by(Experiment, subject, con) %>% 
    summarise(p = mean(accuracy)),
  math %>% 
    filter(Experiment %in% c("1E")) %>% 
    group_by(Experiment, subject) %>% 
    mutate(con = bin_ratings(con, 0, 10, 6)) %>% 
    group_by(Experiment, subject, con) %>% 
    summarise(p = mean(accuracy))
) %>% 
  left_join(select(math_metrics, Experiment, subject, proportion_correct)) %>% 
  ungroup() %>% 
  mutate(Performance = cut2(proportion_correct, g = 3))
res <- tmp %>% 
  filter(con==1) %>% 
  group_by(Performance) %>% 
  group_modify(~tidy(t.test(.$p, mu = 0.25))) %>% 
  glue_data("{round(estimate*100, 1)}%, 95%CI = [{round(conf.low*100, 1)}, {round(conf.high*100, 1)}]")
# tmp %>%   
#   ggplot(aes(con, p, col = Performance)) +
#   scale_y_continuous(
#     "Proportion correct",
#     limits = 0:1
#   ) +
#   labs(x = "Confidence") +
#   stat_summary(geom = "line", aes(group=Performance)) +
#   stat_summary()
```

To begin to address this issue, we took the data of Experiments 1A-F, and asked if performance did indeed drop below 25% at the lowest level of confidence. We did this analysis separately for performance tertiles of participants, because good performing students should necessarily guess less than poor performers. In support of the guessing hypothesis, none of the performance tertiles had performance below 25% at the lowest level of confidence (Low performers: `r res[1]`, medium performers: `r res[2]`, high performers: `r res[3]`). In fact, as can be seen, the lowest performers were almost exactly at 25% at the lowest level of confidence. Nevertheless, these analyses were preliminary, and a stronger test of our hypothesis would be to experimentally contrast tasks where guessing was possible and where it was not. Therefore, we put the idea that resolution metrics are related to domain knowledge in multiple-choice tasks, possibly due to guessing, to an experimental test. In Experiment 2, we directly compared indices of metacognitive resolution, and their relation to domain skill, between free-response trials and multiple-choice trials.

@@

```{r trivia-dtab, results = 'asis'}
tmp1 <- trivia_metrics %>% 
  select(
    session, condition, proportion_correct, 
    mean_rating, gamma_old, zroc_da, zroc_auc_gamma
  ) %>% 
  gather(key, value, -session, -condition) %>% 
  spread(condition, value) %>% 
  mutate(difference = `2AFC`-Free) %>% 
  group_by(key) %>% 
  group_modify(~tidy(t.test(.$difference))) %>% 
  # mutate(p.value = pvalue(p.value), parameter = round(parameter, 0)) %>% 
  mutate_if(is.numeric, ~format(round(., 2), nsmall = 2)) %>%
  ungroup() %>% 
  transmute(
    key = key,
    result = glue("{estimate} [{conf.low}, {conf.high}]")
  ) %>% 
  left_join(new_names) %>% 
  select(-key) %>% 
  spread(Metric, result) %>% 
  add_column(condition = "2AFC - Free", .before = 1)
tmp2 <- trivia_metrics %>%
  # left_join(tmp_mc) %>% 
  gather(
    key, 
    value, 
    proportion_correct, mean_rating, gamma_old, zroc_auc_gamma, zroc_da,
  ) %>% 
  group_by(key, condition) %>%
  group_modify(~ data.frame(t(smean.cl.boot(.x$value, B = 1000)))) %>%
  ungroup() %>%
  left_join(new_names) %>%
  # mutate(Experiment = factor(exp)) %>%
  mutate_at(vars(Mean, Lower, Upper), ~ format(round(., 2), nsmall = 2)) %>%
  mutate(result = glue("{Mean} [{Lower}, {Upper}]")) %>%
  select(condition, Metric, result) %>%
  spread(Metric, result) %>%
  select(condition, `Proportion Correct`, `Mean confidence`, everything())
bind_rows(tmp2, tmp1) %>% 
  kable(
    booktabs = TRUE,
    escape = FALSE,
    caption = "Measures of general information test performance and metacognition in Experiment 2. Numbers indicate means with 95\\% CIs in square brackets."
  ) %>% 
  kable_styling(latex_options = "scale_down")
```

@@

```{r trivia-ctab}
tmp <- trivia_metrics %>%
  gather(key, value, gamma_old, zroc_da, zroc_auc_gamma) %>%
  left_join(new_names) %>%
  group_by(condition, Metric)
tab <- group_modify(
  tmp,
  ~tidy(cor.test(.x$value, .x$proportion_correct, method = "kendall"))
) %>%
  mutate(p.value = pvalue(p.value)) %>%
  mutate(result = format(round(estimate, 2), nsmall = 2, trim = TRUE)) %>%
  mutate(result = glue("{result} ({p.value})")) %>%
  select(condition, Metric, result) %>%
  ungroup() %>%
  pivot_wider(names_from = c(Metric), values_from = result)
kable(
  tab, 
  booktabs = TRUE,
  caption = "Correlation between general information test performance and three metrics of metacognitive resolution in Experiment 2."
  )
```

```{r trivia-rfig, fig.cap = "Scatterplot of participants' resolution metrics and proportions correct across the two conditions in Experiment 2. Lines of best fit and 95\\% CIs are from an ordinary linear regression model.", fig.height = 5}
trivia_metrics %>%
  gather(key, value, zroc_auc_gamma, zroc_da, gamma_old) %>%
  left_join(new_names) %>% 
  ggplot(aes(proportion_correct, value)) +
  geom_hline(yintercept = 0, lty = 2, size = .25) +
  scale_y_continuous(
    "Resolution",
    breaks = pretty_breaks()
  ) +
  scale_x_continuous(
    "Proportion correct",
    breaks = pretty_breaks()
  ) +
  geom_point(shape = 21, alpha = .5, size = .75) +
  geom_smooth(method = "lm", col = "black", size = .4) +
  facet_grid(Metric ~ condition, scales = "free_y") +
  theme(aspect.ratio = 1)
```
